Spatiotemporal Multi-Resolution Approximation of the Amari Type Neural Field Model

### 2012-01-17 Multiresolution_framework/ltx/manuscript.tex
working on MRA SS model section

\section{Multi-Resolution Cortical Architecture}
The human neocortex has an inherent multi-resolution architecture. On finer scales... where any measurable spatiotemporal neural field can be written as 
\begin{equation}
	v\left(\mathbf{r},t\right) = \sum_{j} v_j\left(\mathbf{r},t\right),
\end{equation}
where $v$ is a voltage field that may be measured with a multi-electrode array or voltage sensitive dye for example, $\mathbf{r}\in \mathbb{R}^n$ (where $n\le3$) describes the spatial location with the field, $t$ denotes time, and $j$ indexes the different levels of resolution in the field.

@@

\section{MRA of the IDE in State-Space}
The multi-resolution approximation \cite{Mallat1989a} of the neural IDE (MRAIDE) is obtained by decomposing both the field, $v_t(\cdot)$, and the connectivity kernel, $w(\cdot)$, (assuming square-integrable functions) using translations and dilations of a scaling function $\phi(r)$ and a mother wavelet $\psi(r)$. Considering a one-dimensional field, the connectivity kernel is decomposed as,
\begin{equation}
 w\left(r-r'\right)=\sum_{l \in \mathbb{Z}}\alpha_{j_0,l}\phi_{j_0,l}\left(r-r'\right)+\sum_{j\geq j_0}^{\infty} \sum_{l \in \mathbb{Z}}\beta_{j,l}\psi_{j,l}\left(r-r'\right), 
\label{eq:KernelExpansion}
\end{equation}
where $\alpha_{j_0,l}$ are the approximation coefficients at the lowest scale $j_0$, and $\beta_{j,l}$ are the detail coefficients at different scales $j$, with $\phi_{j,l}\left(r\right)=2^{\frac{j}{2}}\phi\left(2^jr-l\right) $ and $\psi_{j,l}\left(r\right)=2^{\frac{j}{2}}\psi\left(2^jr-l\right)$. Integers $j$ and $l$ are the scale and translation parameters, respectively. Scaling functions retain the lowest frequency components, while the wavelet functions extract successively higher frequency components due to their low-pass and band-pass properties, respectively. The field is decomposed using
\begin{equation}
 v_t\left(r\right)=\sum_{l \in \mathbb{Z}}x_{t,j_{0},l}\phi_{j_{0},l}\left(r\right)+\sum_{j\geq j_0}^{\infty} \sum_{l \in \mathbb{Z}} \check{x}_{t,j,l}\psi_{j,l}\left(r\right),
\label{eq:FieldExpansion}
\end{equation}
where $x_{t,j_{0},l}$ and $\check{x}_{t,j,l} $ are the coefficients of the expansion and constitute the state vector at time $t$. \dean{The decomposition enables a separation of the spatial characteristics, which are considered static functions, that are weighted by dynamic coefficients.  }

Equations \ref{eq:KernelExpansion} and \ref{eq:FieldExpansion} are infinite series expansions and must be truncated at some level $j$. The field and the connectivity kernel are approximated by
\begin{align}
	w\left(r-r'\right) &\approx \boldsymbol\theta^\top\boldsymbol\lambda\left(r-r'\right) 
	\label{eq:KernelFiniteExpansion} \\
	v_t\left(r\right) &\approx \boldsymbol\mu^\top\left(r\right)\mathbf{x}_t,
	\label{eq:FieldFiniteExpansion}
\end{align}
where the unknown parameter and state vectors, $\boldsymbol\theta \in \mathbb{R}^{n_{\theta}}$ and $\mathbf{x}_t \in \mathbb{R}^{n_x}$, are defined as 
\begin{align}
\boldsymbol\theta^\top &=[\begin{array}{ccccc} \boldsymbol\alpha_{j_0}^\top & \boldsymbol\beta_{j_0}^\top & \boldsymbol\beta_{j_0+1}^\top & \cdots & \boldsymbol\beta_{j}^\top \end{array}] 
\label{KernelWeights} \\
\mathbf{x}_{t}^\top &=[\begin{array}{ccccc}\mathbf{x}_{t,j_{0}}^\top &  \check{\mathbf{x}}_{t,j_{0}}^\top & \check{\mathbf{x}}_{t,j_{0}+1}^\top & \cdots & \check{\mathbf{x}}_{t,j}^\top\end{array}].
\label{FieldWeights}
\end{align}
The kernel and field approximation coefficient vectors, $\boldsymbol \alpha_{j_0}$ and $\mathbf{x}_{t,j_{0}}$, contain all the coefficients $\left\lbrace\alpha_{j_0, l}:l \in \mathbb{Z} \right\rbrace $ and $\left\lbrace x_{t,j_0, l}: l \in \mathbb{Z}\right\rbrace$, respectively. Similarly, the kernel and the field detail coefficient vectors, $\boldsymbol\beta_{j}$ and $\check{\mathbf{x}}_{t,j}$, contain all the coefficients $\left\lbrace \beta_{j,l} :l \in \mathbb{Z}\right\rbrace$ and $\left\lbrace  \check x_{t,j, l}:l \in \mathbb{Z}\right\rbrace$, respectively.

The vectors of the kernel and the field scaling and wavelet functions, $\boldsymbol\lambda$ and $\boldsymbol\mu$, respectively, are defined by
\begin{align}
    \label{KernelBasisVector}
    \boldsymbol\lambda^\top(r-r')=&\left[
    \boldsymbol\phi_{j_0}^\top(r-r') ~
    \boldsymbol\psi_{j_0}^\top(r-r') ~ 
    \boldsymbol\psi_{j_0+1}^\top(r-r') ~
    \cdots ~
    \boldsymbol\psi_{j}^\top(r-r')\right] \\
    \label{FieldBasisVector}
    \boldsymbol\mu^\top (r) =& \left[
    \boldsymbol\phi_{j_0}^\top(r) ~ 
    \boldsymbol\psi_{j_0}^\top(r) ~ 
    \boldsymbol\psi_{j_0+1}^\top(r) ~ 
    \cdots ~ 
    \boldsymbol\psi_{j}^\top(r)
\right]   
\end{align}

>>>>

\section{Multi-Resolution Cortical Architecture}
The human neocortex has an inherent multi-resolution architecture. On finer scales... where any measurable spatiotemporal neural field can be written as 
\begin{equation}
	v\left(\mathbf{r},t\right) = \sum_{j} v_j\left(\mathbf{r},t\right),
\end{equation}
where $v\left(\mathbf{r},t\right)$ is a voltage field that may be measured with a multi-electrode array or voltage sensitive dye, $\mathbf{r}\in \mathbb{R}^n$ (where $n\le3$) describes the spatial location with the field, $t$ denotes time, and $j$ indexes the different levels of resolution in the field.

@@

\section{MRA of the IDE in State-Space}
\dean{The purpose of forming the multi-resolution approximation (MRA) of the integro-differential equation (IDE) is to form a multi-resolution finite-dimensional state-space model that best captures the spatiotemporal characteristics of the neural field. Given such a model, it is then possible to perform state and parameter estimation in an efficient and robust manner, thus allowing for subject-specific models.}

The multi-resolution approximation \cite{Mallat1989a} of the neural IDE (MRAIDE) is obtained by decomposing both the field, $v_t(\cdot)$, and the connectivity kernel, $w(\cdot)$, (assuming square-integrable functions) using translations and dilations of a scaling function $\phi(\mathbf{r})$ and a mother wavelet $\psi(\mathbf{r})$. Considering an $n$-dimensional field, the static connectivity kernel is decomposed as,
\begin{equation}
 w\left(\mathbf{r}\right)=\sum_{\mathbf{l} \in \mathbb{Z}^n}\alpha_{j_0,\mathbf{l}} \underbrace{\phi_{j_0,\mathbf{l}}\left(\mathbf{r}\right)}_{\text{scaling}} + \sum_{j\geq j_0}^{\infty} \sum_{\mathbf{l} \in \mathbb{Z}^n}\beta_{j,\mathbf{l}} \underbrace{\psi_{j,\mathbf{l}}\left(\mathbf{r}\right)}_{\text{wavelet}}, 
\label{eq:KernelExpansion}
\end{equation}
where $\alpha_{j_0,\mathbf{l}}$ are the approximation coefficients at the lowest scale $j_0$, and $\beta_{j,\mathbf{l}}$ are the detail coefficients at different scales $j$, with
\begin{align}
\phi_{j,\mathbf{l}}\left(\mathbf{r}\right) =& 2^{\frac{j}{2}}\phi\left(2^j\mathbf{r}-\mathbf{l}\right) \\	
\psi_{j,\mathbf{l}}\left(\mathbf{r}\right) =& 2^{\frac{j}{2}}\psi\left(2^j\mathbf{r}-\mathbf{l}\right).
\end{align}
The parameters $j$ and $\mathbf{l}$ control the scale (dilation) and translation (spatial shift), respectively. \dean{NEED SOMETHING ABOUT HOW j AND l ARE SET HERE - AS IN THAT THEY ARE KNOWN - AND THEAT WE ONLY NEED TO ESTIMATE THE COEFFICIENTS.} Scaling functions retain the lowest frequency components, due to their low-pass properties, while the wavelet functions extract successively higher frequency components due to their band-pass characteristics.

% At each spatial scale indexed by $j$ (level of decomposition), the neural field can be written as
% \begin{equation}
% 	v_{j,t}\left(\mathbf{r}\right) = \begin{cases}
% 	\sum_{l \in \mathbb{Z}}x_{j,l}(t) \underbrace{\phi_{j,l}\left(\mathbf{r}\right)}_{\text{scaling function}} + \check{x}_{t,j,l}\underbrace{\psi_{j,l}\left(\mathbf{r}\right)}_{\text{mother wavelet}} & \text{if $j=0$},\\
% 	\sum_{l \in \mathbb{Z}}\check{x}_{t,j,l}\psi_{j,l}\left(\mathbf{r}\right) & \text{otherwise}
% \end{cases}
% \end{equation}

The dynamic neural field is decomposed using
\begin{equation}
 v_t\left(\mathbf{r}\right)=\sum_{\mathbf{l} \in \mathbb{Z}^n}x_{t,j_{0},\mathbf{l}} \underbrace{\phi_{j_0,\mathbf{l}}\left(\mathbf{r}\right)}_{\text{scaling}} + \sum_{j\geq j_0}^{\infty} \sum_{\mathbf{l} \in \mathbb{Z}^n} \check{x}_{t,j,\mathbf{l}} \underbrace{\psi_{j,\mathbf{l}}\left(\mathbf{r}\right)}_{\text{wavelet}},
\label{eq:FieldExpansion}
\end{equation}
where $x_{t,j_{0},\mathbf{l}}$ and $\check{x}_{t,j,\mathbf{l}} $ are the dynamic coefficients of the expansion, constituting the state vector at time $t$. \dean{The decomposition enables a separation of the static spatial characteristics, which are considered static functions, that are weighted by dynamic coefficients.  }

Equations \ref{eq:KernelExpansion} and \ref{eq:FieldExpansion} are infinite series expansions and must be truncated to some level \dean{$J$} in order to solve the estimation problem. In other words, the neural field must be band-limited up to a given spatial frequency. Assuming the field and connectivity kernel are spatially band-limited they can be accurately approximated by
\begin{align}
	w\left(\mathbf{r}\right) &\approx \boldsymbol\theta^\top\boldsymbol\lambda\left(\mathbf{r}\right) 
	\label{eq:KernelFiniteExpansion} \\
	v_t\left(\mathbf{r}\right) &\approx \boldsymbol\mu^\top\left(\mathbf{r}\right)\mathbf{x}_t,
	\label{eq:FieldFiniteExpansion}
\end{align}
where the unknown parameter and state vectors, $\boldsymbol\theta \in \mathbb{R}^{n_{\theta}}$ and $\mathbf{x}_t \in \mathbb{R}^{n_x}$, are defined as 
\begin{align}
\boldsymbol\theta^\top &=[\begin{array}{ccccc} \boldsymbol\alpha_{j_0}^\top & \boldsymbol\beta_{j_0}^\top & \boldsymbol\beta_{j_0+1}^\top & \cdots & \boldsymbol\beta_{J}^\top \end{array}] 
\label{KernelWeights} \\
\mathbf{x}_{t}^\top &=[\begin{array}{ccccc}\mathbf{x}_{t,j_{0}}^\top &  \check{\mathbf{x}}_{t,j_{0}}^\top & \check{\mathbf{x}}_{t,j_{0}+1}^\top & \cdots & \check{\mathbf{x}}_{t,J}^\top\end{array}].
\label{FieldWeights}
\end{align}
The kernel and field approximation coefficient vectors, $\boldsymbol \alpha_{j_0}$ and $\mathbf{x}_{t,j_{0}}$, contain all the coefficients $\left\lbrace\alpha_{j_0, \mathbf{l}}:\mathbf{l} \in \mathbb{Z}^n \right\rbrace $ and $\left\lbrace x_{t,j_0, \mathbf{l}}: \mathbf{l} \in \mathbb{Z}^n\right\rbrace$, respectively. Similarly, the kernel and the field detail coefficient vectors, $\boldsymbol\beta_{j}$ and $\check{\mathbf{x}}_{t,j}$, contain all the coefficients $\left\lbrace \beta_{j,\mathbf{l}} :\mathbf{l} \in \mathbb{Z}^n\right\rbrace$ and $\left\lbrace  \check x_{t,j,\mathbf{l}}:\mathbf{l} \in \mathbb{Z}^n\right\rbrace$, respectively.

The vectors of the kernel and the field scaling and wavelet functions, $\boldsymbol\lambda$ and $\boldsymbol\mu$, respectively, are defined by
\begin{align}
    \label{KernelBasisVector}
    \boldsymbol\lambda^\top(\mathbf{r})=&\left[
    \begin{array}{ccccc}\boldsymbol\phi_{j_0}^\top(\mathbf{r}) &
    \boldsymbol\psi_{j_0}^\top(\mathbf{r}) & 
    \boldsymbol\psi_{j_0+1}^\top(\mathbf{r}) &
    \cdots &
    \boldsymbol\psi_{J}^\top(\mathbf{r})\end{array}\right] \\
    \label{FieldBasisVector}
    \boldsymbol\mu^\top (\mathbf{r}) =&\left[
    \begin{array}{ccccc}\boldsymbol\phi_{j_0}^\top(\mathbf{r}) &
    \boldsymbol\psi_{j_0}^\top(\mathbf{r}) & 
    \boldsymbol\psi_{j_0+1}^\top(\mathbf{r}) &
    \cdots &
    \boldsymbol\psi_{J}^\top(\mathbf{r})\end{array}\right]
\end{align}
